---
title: "Getting Started with paddle"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting_started_with_paddleR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r eval=FALSE}
library(paddleR)
```

## Overview

The `paddleR` package provides a complete R interface to the [Paddle](https://www.paddle.com/) Billing [API](https://developer.paddle.com/api-reference/overview), enabling you to manage customers, subscriptions, transactions, invoices, and more directly from R.

Whether you're building internal dashboards, automated billing workflows, or a SaaS application backend, paddle makes it easy to interact with Paddle's RESTful endpoints in a consistent and validated way.

## Setting the API Mode

`PaddleR` supports two environments:

* Live (https://api.paddle.com) for production
* Sandbox (https://sandbox-api.paddle.com) for safe testing

By default, the package uses live mode. You can switch modes using:

```{r eval=FALSE}
# Set sandbox mode for testing
set_paddle_mode("sandbox")

# Revert back to live mode
set_paddle_mode("live")
```

You can check the current mode and base URL with:

```{r eval=FALSE}
get_paddle_mode()  # Returns "live" or "sandbox"
get_paddle_url()   # Returns full API base URL
```

## Setting Your API Key

To authenticate with the Paddle API, you must set your API key(s) as environment variables. You have two options:

### Option 1: Single Key (Live OR Sandbox)

```{r}
Sys.setenv(PADDLE_KEY = "your-live-or-sandbox-key")
```

### Option 2: Recommended – Separate Keys for Live and Sandbox

```{r eval=FALSE}
Sys.setenv(PADDLE_KEY_LIVE = "your-live-key")
Sys.setenv(PADDLE_KEY_SANDBOX = "your-sandbox-key")
```

This allows the package to automatically choose the correct key based on the mode selected with `set_paddle_mode()`. Please make sure your API has all the necessary permissions for the operations you want to perform.

### Example: Initialize and Fetch Data

```{r eval=FALSE}
# Set mode to sandbox for testing
set_paddle_mode("sandbox")

# List products, subscriptions, or customers
products <- paddle_list_products()
subscriptions <- paddle_list_subscriptions()
```

## Package Initialization Logic

Internally, the package uses a hidden environment to store the current mode and URL:

```{r eval=FALSE}
.paddle_env <- new.env(parent = emptyenv())
.paddle_env$mode     <- "live"
.paddle_env$base_url <- "https://api.paddle.com"
```

These are updated dynamically using the exported helpers:

* set_paddle_mode("sandbox")
* get_paddle_mode()
* get_paddle_url()

## Next Steps
Now that you have paddleR set up, you can start building your Paddle integration:

* Use `paddle_list_*()` functions to explore available resources like products, subscriptions, customers, etc.
* Create and manage customers with `paddle_create_customer()`, `paddle_update_customer()`, etc.
* Manage subscriptions using `paddle_create_subscription()`, `paddle_update_subscription()`, etc.
* Handle transactions with `paddle_list_transactions()`, `paddle_get_transaction()`, etc.
* Work with invoices using `paddle_list_invoices()`, `paddle_get_invoice()`, etc.
* Manage discounts with `paddle_list_discounts()`, `paddle_create_discount()`, etc.
* Use `paddle_list_prices()` to explore available prices for products
* Create and manage products with `paddle_create_product()`, `paddle_update_product()`, etc.

## Need Help?

* File an issue on GitHub if something doesn’t work
* Some of API endpoints may not be implemented yet, so feel free to request new features
* Check the [Paddle API documentation](https://developer.paddle.com/api-reference/overview) for more details on available endpoints
